\documentclass{aa}

\usepackage{showyourwork}
\usepackage{hyperref}
\hypersetup{
  colorlinks,
  citecolor=cyan,
  linkcolor=red,
  urlcolor=blue
}
\usepackage{graphicx}
\usepackage{txfonts}
\usepackage{orcidlink}
\usepackage{siunitx}
% common astronomical symbols
\let\sun\odot
\DeclareSIUnit\erg{erg}
\DeclareSIUnit\jansky{Jy}
\DeclareSIUnit\year{yr}
\DeclareSIUnit\solarluminosity{\ensuremath{L_\sun}}
\DeclareSIUnit\solarmass{\ensuremath{M_\sun}}
\DeclareSIUnit\solarradius{\ensuremath{R_\sun}}
\DeclareSIUnit\arcsecond{as}
\DeclareSIUnit\astronomicalunit{au}

\defcitealias{2019MNRAS.487..691P}{Paper~I}
\defcitealias{2021MNRAS.508.1675E}{Paper~II}
\defcitealias{2021MNRAS.508.3611P}{Paper~III}

\begin{document}

   \title{Disks in transition}

   \subtitle{I. Internal photoevaporation and disk dispersal}

   \author{Giovanni Picogna \orcidlink{0000-0003-3754-1639}}

   \institute{Universit채ts-Sternwarte, Ludwig-Maximilians-Universit채t M체nchen, Scheinerstr. 1, D-81679 M체nchen, Germany\\
   \email{picogna@usm.lmu.de}
    }

   \date{Received \today; accepted \today}

   \abstract
      {The strong X-ray irradiation from young solar-type stars may play a crucial role in the thermodynamics and chemistry of circumstellar discs, driving their evolution in the last stages of disc dispersal when the accretion rates onto the central star drop and thermal winds become the only viable way to dissipate the disc. Yet, we have discovered a large population of disks with cavities that are not in transition, and no concrete evidence of a disc in his last evolutionary stages.}
      {We study the influence of the cavity sizes on the disc mass-loss rates due to X-ray irradiation for stars with masses between $0.1$ and \SI{1}{\solarmass} and derive observables to be compared with the transition discs population.}
      {We ran a series of radiative-hydrodynamical simulations to derive the equilibrium wind mass-loss rate as a function of cylindrical radius from the central star for varying stellar and disc properties, and apply the resulting profiles to 1D long term disc evolution. We compared then the synthetic transition disc population with the observed one.}
      {We  found a simple scaling between cavity size, stellar mass and disc mass that can be readily applied to 1D viscous evolution codes to model the final disc dispersal in a proper fashion.}
      {We provide new analytical relations for the mass-loss rates and profiles of photoevaporative winds as a function of the stellar mass that can be used in disc and planet population synthesis models. Our photoevaporative models correctly predict the observed trend of inner-disc lifetime as a function of stellar mass with an increased steepness for stars smaller than \SI{0.3}{\solarmass}, indicating that X-ray photoevaporation is a good candidate to explain the observed disc dispersal process.}

   \keywords{}

   \maketitle

   \section{Introduction} \label{sec:intro}

   \input{introduction}

   \section{Dependence on the cavity size}\label{sec:cavity}

   \section{Methods}\label{sec:methods}

      \subsection{Radiative transfer}

      \subsection{Hydrodynamical model}\label{sec:hydro-model}

      \subsection{Viscous evolution}

   \section{Results}\label{sec:results}

      \subsection{Disc profiles}

      \subsection{Mass-loss rates}\label{sec:mdot}

      \subsection{Wind profiles}\label{sec:wind-prof}

      \subsection{Gas cavity radius dependence}\label{sec:gap-dependance}

   \section{Discussion}\label{sec:discussion}

      \subsection{Comparison with observations}

      \subsection{Comparison with other models}

   \section{Conclusions}\label{sec:conclusions}

   \section{Data availability}
      The data underlying this article and the scripts used to create the Figures are available at \ldots

   \begin{appendix}

   \end{appendix}

   \begin{acknowledgements}
      GP acknowledges support from the DFG Research Unit 'Transition Disks' (FOR 2634/2).
      This work was performed on the computing facilities of the Computational Center for Particle and Astrophysics (C2PAP).
      GP would like to thank Kristina Monsch and Tommaso Grassi for their insightful comments on the manuscript and during the development of the project.
      This research was supported by the Excellence Cluster ORIGINS which is funded by the Deutsche Forschungsgemeinschaft (DFG,  German Research Foundation) under Germany's Excellence Strategy - EXC-2094 - 390783311.
      This paper utilizes the D'Alessio Irradiated Accretion Disk (DIAD) code. We wish to recognize the work of Paola D'Alessio, who passed away in 2013. Her legacy and pioneering work live on through her substantial contributions to the field.
   \end{acknowledgements}

   \bibliographystyle{aa} % style aa.bst
   \bibliography{bib}

\end{document}
